Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-08-13 paper/current/main.tex
merge paragraphs

How far apart are two places?
The question is surprisingly hard to answer when it involves human mobility.
Although geographic distance has been constraining  human movements throughout history, it is becoming less relevant in a world connected by rapid transit and global airline networks.
For instance, a person living in Australia is more likely to migrate to the United Kingdom, a far-away country with similar language and culture, than to a much closer country such as Indonesia~\autocite{pew2018migration}.
Similarly, a student in South Korea is more likely to attend a university in Canada than one in North Korea~\autocite{unesco2019students}.
Although geographic distance has been used as the most prominent basis for models of mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation~\autocite{simini2012universal} models, there have been attempts to define alternative notions of distance, or ~\textit{functional distances}~\autocite{boschma2005proximity, brown1970functional, brown1970migration, kim2018functional} from real-world data or \textit{a priori} relationships between geographic entities.
Yet, functional distances are often low-resolution, at the level of countries rather than regions, cities, or organizations, and rely on somewhat arbitrary decisions about which factors are most relevant.
Moreover, functional distances have focused on only a single facet of mobility at a time, whereas real-world mobility is multi-faceted, influenced simultaneously by geography, language, culture, history, and economic opportunity.
To capture such complexity of human mobility, new approaches that incorporate its many facets into a single high-resolution functional distance are needed.

Here, we propose using a neural embedding model that draws from real-world mobility trajectories can encode the complex landscape of human mobility into a dense, continuous, and meaningful vector-space representation, from which not only a meaningful \textit{functional distance} between places but also relationship based on culture, language, and even prestige along with geographic relationship can be probed.

Scientific mobility is a central driver of the increasingly-globalized scientific enterprise~\autocite{czaika2018globalisation, altbach2004globalization, box2008competition}, being strongly related to innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}.
Moreover, scientific mobility has known structure, such as the centrality of the United States and the United Kingdom, and the strong prestige hierarchy~\autocite{auriol2010careers, clauset2015hierarchy, deville2014career} that should be captured in a representation.
However, understandings of scientific mobility have been limited by the sheer scope and complexity of the phenomenon~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics}, being further confounded by the diminishing role of geography in shaping the landscape of scientific mobility.

>>>>

How far apart are two places?
The question is surprisingly hard to answer when it involves human mobility.
Although geographic distance has been constraining human movements throughout history, it is becoming less relevant in a world connected by rapid transit and global airline networks.
For instance, a person living in Australia is more likely to migrate to the United Kingdom, a far-away country with similar language and culture, than to a much closer country such as Indonesia~\autocite{pew2018migration}.
Similarly, a student in South Korea is more likely to attend a university in Canada than one in North Korea~\autocite{unesco2019students}.
Although geographic distance has been used as the most prominent basis for models of mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation~\autocite{simini2012universal} models, there have been attempts to define alternative notions of distance, or ~\textit{functional distances}~\autocite{boschma2005proximity, brown1970functional, brown1970migration, kim2018functional} from real-world data or \textit{a priori} relationships between geographic entities.

Yet, functional distances are often low-resolution, at the level of countries rather than regions, cities, or organizations, and have focused on only a specific single facet of mobility at a time, whereas real-world mobility is multi-faceted, influenced simultaneously by geography, language, culture, history, and economic opportunity.  Furthermore, even though the network has been using a powerful unifying approach for understanding human mobility, discrete characteristic of the network makes hard to apply simple algebraic operations or common algorithms directly, which can be easily solved with a fine-grained representation of each entity. To capture such complexity of human mobility and overcome these limitations, new approaches that can produce an coherent representation and incorporate its many facets into a single high-resolution functional distance are needed.

Here, we propose using a neural embedding model that draws from real-world mobility trajectories can encode the complex landscape of human mobility into a dense, continuous, and meaningful vector-space representation, from which not only a meaningful \textit{functional distance} between places but also relationship based on culture, language, and even prestige along with geographic relationship can be probed. We leverage a massive dataset that documents scientific mobility captured in bibliometric records. Scientific mobility, which is central driver of the increasingly-globalized scientific enterprise~\autocite{czaika2018globalisation, altbach2004globalization, box2008competition}, being strongly related to innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation}, impact~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage}, collaboration~\autocite{rodrigues2016mobility}, and the diffusion of knowledge~\autocite{braunerhjelm2020labor, azoulay2011diffusion, morgan2018prestige}, is not only an important topic on science of science  but also ideal for our study thanks to its richness and well-known structural properties such as the centrality to scientific-advance country, and the strong prestige hierarchy~\autocite{auriol2010careers, clauset2015hierarchy, deville2014career}. In spite of the importance of the problem, understandings of scientific mobility have been limited by the sheer scope and complexity of the phenomenon~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics}, being further confounded by the diminishing role of geography in shaping the landscape of scientific mobility.
